---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# plfs

The goal of plfs is to provide a long dataset of the Periodic Labour Force Survey (PLFS) from India.

## Example

Install the package from GitHub and load it:

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("rahulsh97/plfs")
```

```{r load}
library(plfs)
```

Because of the datasets size, the package provides a function to download the datasets and create a local DuckDB
database. This results in a CRAN-compliant package.

Here is how to get the plfs database ready for use:

```{r download, eval=FALSE}
plfs_download()
```

TODO: Add example similar to ASI.

# Adding older/newer years

Be sure to use the yearly survey (e.g., Jul 23 - Jul 24)

1. Install the Nesstar Explorer (e.g. plfs 2023-24 includes it)
2. Extract the RAR files downloaded from the microdata website to data-raw/202223 or what year you are adding
3. Export the .Nesstar file to Stata (SAV) format with "Export Datasets" and the metadata with "Export DDI" using the Nesstar Explorer
4. Update `00-tidy-data.r` and run it
5. Update the available datasets in `R/available_datasets.R`
6. Update the new RDS files in the 'Releases' section of the GitHub repository
7. Regenerate the database with `plfs_delete()` and `plfs_download()`
